Realizability-Preserving DG-IMEX Method for the Two-Moment Model of Fermion Transport \tnoteref{support}

### 2018-06-20 Documents/M1/realizableFermionicM1/sections/DGIMEXscheme.tex
PARSD -> PD-ARS

\begin{rem}
  For IMEX scheme PARSD (\ref{app:PARSD}), which does not invoke the correction step (i.e., $\alpha=0$), $\vect{\cM}_{\bK}^{n+1}=\vect{\cM}_{\bK}^{(s)}$.  
\end{rem}

>>>>

\begin{rem}
  For IMEX scheme PD-ARS (\ref{app:PD-ARS}), which does not invoke the correction step (i.e., $\alpha=0$), $\vect{\cM}_{\bK}^{n+1}=\vect{\cM}_{\bK}^{(s)}$.  
\end{rem}

### 2018-06-20 Documents/M1/realizableFermionicM1/sections/appendix.tex
PARSD -> PD-ARS

\paragraph{IMEX PARSD}

We have found diffusion accurate, positivity-preserving IMEX schemes of type ARS that are second-order accurate in the streaming limit, which we refer to as IMEX PARSD; see \ref{app:PARSD}.  
For these schemes, $c_{\mbox{\tiny Sch}}= 1 - 2\epsilon$ with $\epsilon \in [0, 1/2)$.
Here we give an example by setting $\epsilon=0.1$.
\begin{align*}
  \tilde{a}_{21} & = 1.0, \\
  \tilde{a}_{31} & = 0.5, \quad \tilde{a}_{32} = 0.5, \\
  a_{22} & = 1.0, \nonumber \\
  a_{32} & = 0.4 \,( = 0.5 - \epsilon\,), \quad a_{33} = 0.6 \,( = 0.5 + \epsilon\,). 
\end{align*}
This scheme is GSA, $\alpha=0$, and requires only two implicit solves per time step.

@@

\newpage 
\section{Construction of PARSD}
\label{app:PARSD}

It can be shown that no 2-stage positivity-preserving IMEX exist \cite{hu_etal_2018}.
Hence we considered 3-stage IMEX scheme.
There are two kinds of implicit DIRK scheme, Type A and Type ARS.
Here we discuss the simpler case, Type ARS, first.
The Butcher tableau for ARS type IMEX scheme is
\begin{equation}
  \begin{array}{c | c c c}
  	         0           & 0                    & 0                    & 0                    \\
  	\tilde{c}_{2} & \tilde{a}_{21}       & 0                    & 0                    \\
  	\tilde{c}_{3} & \tilde{a}_{31}       & \tilde{a}_{32}       & 0                    \\ \hline
  	                     & \tilde{w}_{1} & \tilde{w}_{2} & \tilde{w}_{3}
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0            & 0            & 0            \\
  	c_{2} & 0            & a_{22}       & 0            \\
  	c_{3} & 0            & a_{32}       & a_{33}       \\ \hline
  	             & w_{1} & w_{2} & w_{3}
  \end{array}.
\end{equation}
The properties a desired IMEX scheme should have are: 2nd-order accurate in the streaming, globally stiffly accurate, positivity-preserving, diffusion accurate and an affordable time-step restriction.
Then the problem that find such a scheme can be rewritten as an optimization problem:\\
Find $(\tilde{a}_{21}, \tilde{a}_{31}, \tilde{a}_{32}, \tilde{w}_{1}, \tilde{w}_{2}, \tilde{w}_{3}, a_{22}, a_{32}, a_{33}, w_{1}, w_{2}, w_{3})$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ \dfrac{c_{20}}{\tilde{c}_{20}}, \dfrac{c_{30}}{\tilde{c}_{30}}, \dfrac{c_{32}}{\tilde{c}_{32}}\}\\
 \text{under these constraints:} \qquad & \\
 \text{(GSA)} \qquad\ &\tilde{w}_{i} = \tilde{a}_{3i}, \quad w_{i} = a_{3i}, ~\text{for} ~ i = 1,2,3\\
 \text{(consistency)} \qquad\ & \sum_{i=1}^{3}  \tilde{w}_{i} = 1, \quad \sum_{i=1}^{3} w_{i} = 1\\ 
 \text{(2nd-order streaming accuracy)} \qquad\ & \sum_{i=1}^{3}  \tilde{w}_{i}\tilde{c}_{i} = 1/2, \\
 \text{(positivity-preserving)} \qquad & a_{ii} > 0, c_{i0} \geq 0, \tilde{c}_{i0} \geq 0, ~\text{for} ~ i = 2,3 \\
 & c_{32} \geq 0, \tilde{c}_{32} \geq 0, \\
 \text{(diffusion accurate)} \qquad & e^{T}_iA^{-1}\tilde{A}e = 1, ~\text{for} ~ i = 2,3, \\
 \text{(affordable)} \qquad & c_{sch} > 0.
\end{align*}
Imposing those equality constraints (including all the constraints exclude those from the positivity-preserving and affordable requirements), the Butcher tableau can be simplified as:
\begin{equation}
  \begin{array}{c | c c c}
  	     0       & 0            & 0 & 0 \\
  	\frac{1}{2x} & \frac{1}{2x} & 0 & 0 \\
  	     1       & 1-x          & x & 0 \\ \hline
  	             & 1-x          & x & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0 & 0            & 0 \\
  	\frac{1}{2x} & 0 & \frac{1}{2x} & 0 \\
  	     1       & 0 & 1-y          & y \\ \hline
  	             & 0 & 1-y          & y
  \end{array}.
\end{equation}
And the rest inequalities and the problem were left to be: \\
Find $(x,y)$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ 2x, \dfrac{1-2x(1-y)}{y-x}, 2(1-y)\}\\
 \text{under these constraints:} \qquad & x > 0, ~ 1 \geq y > 0, \\
 & x \leq y, ~\dfrac{1}{2x} \geq 1-y.
\end{align*}
A solution family is $x=\frac{1}{2}$, $y=\frac{1}{2} + \epsilon$ with $\epsilon \in [0, \frac{1}{2})$ and $c_{sch} = 1 - 2\epsilon$.
The PARSD scheme follows.
It's clear that $\epsilon = 0$ and $c_{sch} = 1$ is the most optimized solution.

>>>>

\paragraph{IMEX PD-ARS}

We have found diffusion accurate, positivity-preserving IMEX schemes of type ARS that are second-order accurate in the streaming limit, which we refer to as IMEX PD-ARS; see \ref{app:PD-ARS}.  
For these schemes, $c_{\mbox{\tiny Sch}}= 1 - 2\epsilon$ with $\epsilon \in [0, 1/2)$.
Here we give an example by setting $\epsilon=0.1$.
\begin{align*}
  \tilde{a}_{21} & = 1.0, \\
  \tilde{a}_{31} & = 0.5, \quad \tilde{a}_{32} = 0.5, \\
  a_{22} & = 1.0, \nonumber \\
  a_{32} & = 0.4 \,( = 0.5 - \epsilon\,), \quad a_{33} = 0.6 \,( = 0.5 + \epsilon\,). 
\end{align*}
This scheme is GSA, $\alpha=0$, and requires only two implicit solves per time step.

@@

\newpage 
\section{Construction of PD-ARS}
\label{app:PD-ARS}

It can be shown that no 2-stage positivity-preserving IMEX exist \cite{hu_etal_2018}.
Hence we considered 3-stage IMEX scheme.
There are two kinds of implicit DIRK scheme, Type A and Type ARS.
Here we discuss the simpler case, Type ARS, first.
The Butcher tableau for ARS type IMEX scheme is
\begin{equation}
  \begin{array}{c | c c c}
  	         0           & 0                    & 0                    & 0                    \\
  	\tilde{c}_{2} & \tilde{a}_{21}       & 0                    & 0                    \\
  	\tilde{c}_{3} & \tilde{a}_{31}       & \tilde{a}_{32}       & 0                    \\ \hline
  	                     & \tilde{w}_{1} & \tilde{w}_{2} & \tilde{w}_{3}
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0            & 0            & 0            \\
  	c_{2} & 0            & a_{22}       & 0            \\
  	c_{3} & 0            & a_{32}       & a_{33}       \\ \hline
  	             & w_{1} & w_{2} & w_{3}
  \end{array}.
\end{equation}
The properties a desired IMEX scheme should have are: 2nd-order accurate in the streaming, globally stiffly accurate, positivity-preserving, diffusion accurate and an affordable time-step restriction.
Then the problem that find such a scheme can be rewritten as an optimization problem:\\
Find $(\tilde{a}_{21}, \tilde{a}_{31}, \tilde{a}_{32}, \tilde{w}_{1}, \tilde{w}_{2}, \tilde{w}_{3}, a_{22}, a_{32}, a_{33}, w_{1}, w_{2}, w_{3})$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ \dfrac{c_{20}}{\tilde{c}_{20}}, \dfrac{c_{30}}{\tilde{c}_{30}}, \dfrac{c_{32}}{\tilde{c}_{32}}\}\\
 \text{under these constraints:} \qquad & \\
 \text{(GSA)} \qquad\ &\tilde{w}_{i} = \tilde{a}_{3i}, \quad w_{i} = a_{3i}, ~\text{for} ~ i = 1,2,3\\
 \text{(consistency)} \qquad\ & \sum_{i=1}^{3}  \tilde{w}_{i} = 1, \quad \sum_{i=1}^{3} w_{i} = 1\\ 
 \text{(2nd-order streaming accuracy)} \qquad\ & \sum_{i=1}^{3}  \tilde{w}_{i}\tilde{c}_{i} = 1/2, \\
 \text{(positivity-preserving)} \qquad & a_{ii} > 0, c_{i0} \geq 0, \tilde{c}_{i0} \geq 0, ~\text{for} ~ i = 2,3 \\
 & c_{32} \geq 0, \tilde{c}_{32} \geq 0, \\
 \text{(diffusion accurate)} \qquad & e^{T}_iA^{-1}\tilde{A}e = 1, ~\text{for} ~ i = 2,3, \\
 \text{(affordable)} \qquad & c_{sch} > 0.
\end{align*}
Imposing those equality constraints (including all the constraints exclude those from the positivity-preserving and affordable requirements), the Butcher tableau can be simplified as:
\begin{equation}
  \begin{array}{c | c c c}
  	     0       & 0            & 0 & 0 \\
  	\frac{1}{2x} & \frac{1}{2x} & 0 & 0 \\
  	     1       & 1-x          & x & 0 \\ \hline
  	             & 1-x          & x & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0 & 0            & 0 \\
  	\frac{1}{2x} & 0 & \frac{1}{2x} & 0 \\
  	     1       & 0 & 1-y          & y \\ \hline
  	             & 0 & 1-y          & y
  \end{array}.
\end{equation}
And the rest inequalities and the problem were left to be: \\
Find $(x,y)$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ 2x, \dfrac{1-2x(1-y)}{y-x}, 2(1-y)\}\\
 \text{under these constraints:} \qquad & x > 0, ~ 1 \geq y > 0, \\
 & x \leq y, ~\dfrac{1}{2x} \geq 1-y.
\end{align*}
A solution family is $x=\frac{1}{2}$, $y=\frac{1}{2} + \epsilon$ with $\epsilon \in [0, \frac{1}{2})$ and $c_{sch} = 1 - 2\epsilon$.
The PD-ARS scheme follows.
It's clear that $\epsilon = 0$ and $c_{sch} = 1$ is the most optimized solution.

### 2018-06-20 Documents/M1/realizableFermionicM1/sections/imex.tex
PARSD -> PD-ARS

Fortunately, IMEX schemes satisfying these properties are easy to find, and we provide an example with $s=3$ (two implicit solves) in \ref{app:butcherTables} (scheme PARSD; see \ref{app:PARSD} for details).  

@@

In Section~\ref{sec:numerical}, we show numerically that the accuracy of scheme PARSD is comparable to the accuracy of scheme PC2.

@@

Here we analyze the absolute stability of the proposed IMEX schemes, PA2+ and PARSD (given in \ref{app:butcherTables}), following \cite{hu_etal_2018}.  
As is commonly done, we do this in the context of the linear scalar equation
\begin{equation}
  \dot{u}=\lambda_{1}\,u+\lambda_{2}\,u,
  \label{eq:scalarODE}
\end{equation}
where $\lambda_{1}\in\mathbb{C}$ and real $\lambda_{2}\le0$.  
On the right-hand side of Eq.~\eqref{eq:scalarODE}, the first (oscillatory) term is treated explicitly, while the second (damping) term is treated implicitly.  
The IMEX schemes can then be written as $u^{n+1} =P(z_{1},z_{2})\,u^{n}$, where $P(z_{1},z_{2})$ is the amplification factor of the scheme, $z_{1}=\dt\,\lambda_{1} = x +iy$, and $z_{2} = \dt\,\lambda_{2} < 0$.  
Stability of the IMEX scheme requires $|P(z_1,z_2)|\leq 1$.  
The stability regions of PA2+ and PARSD are plotted in Figure~\ref{fig:AbsoluteStability}.  
As can be seen from the figures, the absolute stability region of both schemes increases with increasing $|z_{2}|$.  
For a given $|z_{2}|>0$, the stability region of PARSD is larger than that of PA2+.  
In the linear case, a time step that satisfies the absolute stability for the explicit part of the IMEX scheme ($|z_2| = 0$) fulfills the stability requirement for the IMEX scheme as whole ($|z_2| \geq 0$).  
This conclusion is valid for all PARSD schemes with $\epsilon\in [0,0.5)$.
\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPARSD}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPA2+}
  \end{tabular}
   \caption{Boundary of the absolute stability region for different values of $z_2$ for PARSD with $\epsilon = 0.1$ (left panel) and PA2+ (right panel).}
  \label{fig:AbsoluteStability}
\end{figure}

>>>>

Fortunately, IMEX schemes satisfying these properties are easy to find, and we provide an example with $s=3$ (two implicit solves) in \ref{app:butcherTables} (scheme PD-ARS; see \ref{app:PD-ARS} for details).  

@@

In Section~\ref{sec:numerical}, we show numerically that the accuracy of scheme PD-ARS is comparable to the accuracy of scheme PC2.

@@

Here we analyze the absolute stability of the proposed IMEX schemes, PA2+ and PD-ARS (given in \ref{app:butcherTables}), following \cite{hu_etal_2018}.  
As is commonly done, we do this in the context of the linear scalar equation
\begin{equation}
  \dot{u}=\lambda_{1}\,u+\lambda_{2}\,u,
  \label{eq:scalarODE}
\end{equation}
where $\lambda_{1}\in\mathbb{C}$ and real $\lambda_{2}\le0$.  
On the right-hand side of Eq.~\eqref{eq:scalarODE}, the first (oscillatory) term is treated explicitly, while the second (damping) term is treated implicitly.  
The IMEX schemes can then be written as $u^{n+1} =P(z_{1},z_{2})\,u^{n}$, where $P(z_{1},z_{2})$ is the amplification factor of the scheme, $z_{1}=\dt\,\lambda_{1} = x +iy$, and $z_{2} = \dt\,\lambda_{2} < 0$.  
Stability of the IMEX scheme requires $|P(z_1,z_2)|\leq 1$.  
The stability regions of PA2+ and PD-ARS are plotted in Figure~\ref{fig:AbsoluteStability}.  
As can be seen from the figures, the absolute stability region of both schemes increases with increasing $|z_{2}|$.  
For a given $|z_{2}|>0$, the stability region of PD-ARS is larger than that of PA2+.  
In the linear case, a time step that satisfies the absolute stability for the explicit part of the IMEX scheme ($|z_2| = 0$) fulfills the stability requirement for the IMEX scheme as whole ($|z_2| \geq 0$).  
This conclusion is valid for all PD-ARS schemes with $\epsilon\in [0,0.5)$.
\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPARSD}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPA2+}
  \end{tabular}
   \caption{Boundary of the absolute stability region for different values of $z_2$ for PD-ARS with $\epsilon = 0.1$ (left panel) and PA2+ (right panel).}
  \label{fig:AbsoluteStability}
\end{figure}

### 2018-06-20 Documents/M1/realizableFermionicM1/sections/numerical.tex
PARSD -> PD-ARS

To compare the accuracy of the IMEX schemes, we present results from smooth problems in streaming, absorption, and scattering dominated regimes in one spatial dimension.  
For all tests in this subsection, we use third order accurate spatial discretization (polynomials of degree $k=2$) and we employ the maximum entropy closure in the low occupancy limit (i.e., the Minerbo closure).  
We compare results obtained using IMEX schemes proposed here (PA2+ and PARSD) with IMEX schemes from Hu et al. \cite{hu_etal_2018} (PA2), McClarren et al. \cite{mcclarren_etal_2008} (PC2), Pareschi \& Russo \cite{pareschiRusso_2005} (SSP2332), and Cavaglieri \& Bewley \cite{cavaglieriBewley2015} (RKCB2).  
In the streaming test, we also include results obtained with second-order and third-order accurate explicit strong stability-preserving Runge-Kutta methods \cite{gottlieb_etal_2001} in the comparison (SSPRK2 and SSPRK3, respectively).  
See \ref{app:butcherTables} for further details.  
The time step is set to $\dt=0.1\times\dx$.

@@

In Figure~\ref{fig:SineWaveStreaming}, the absolute error for the number density $E_{\mbox{\tiny Abs}}^{1}[\cJ_{h}](t=10)$ is plotted versus $N$ (see figure caption for details).  
Errors obtained with SSPRK3 are smallest and decrease as $N^{-3}$ (cf. bottom black dash-dot reference line), as expected for a scheme combining third-order accurate time stepping with third-order accurate spatial discretization.  
For all the other schemes, using second-order accurate explicit time stepping, the error decreases as $N^{-2}$.  
Among the second-order accurate methods, SSP2332 has the smallest error, followed by RKCB2.  
Errors for the remaining schemes (including SSPRK2) are indistinguishable on the plot.  
\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/SineWaveStreaming}
   \caption{Absolute error (cf. Eq.~\eqref{eq:errorNormAbsolute}) versus number of elements $N$ for the streaming sine wave test.  Results employing various time stepping schemes are compared: SSPRK2 (cyan triangles pointing up), SSPRK3 (cyan triangles pointing down), PA2 (red), PA2+ (purple), PC2 (blue), RKCB2 (dark green), SSP2332 (green), and PARSD (light red circles).  Black dash-dot reference lines are proportional to $N^{-1}$ (top), $N^{-2}$ (middle), and $N^{-3}$ (bottom), respectively.}
  \label{fig:SineWaveStreaming}
\end{figure}

@@

Figure~\ref{fig:SineWaveDamping} shows convergence results, obtained using different values of $\sigma_{\Ab}$, for various IMEX schemes at $t=t_{\mbox{\tiny end}}$.  
Results for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively (see figure caption for further details).  
All the second-order accurate schemes (PA2, PA2+, RKCB2, and SSP2332) display second-order accurate convergence rates (cf. bottom, black dash-dot reference line).  
For $\sigma_{\Ab}=0.1$, SSP2332 is the most accurate among these schemes, while PA2+ is the most accurate for $\sigma_{\Ab}=10$.  
On the other hand, PC2 and PARSD are indistinguishable and display at most first-order accurate convergence, as expected.  
(For $\sigma_{\Ab}=0.1$, PC2 and PARSD are the most accurate schemes for $N=8$ and $N=16$.)

\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/SineWaveDamping}
   \caption{Relative error (cf. Eq.~\eqref{eq:errorNormRelative}) versus number of elements for the damping sine wave test.  Results for different values of the absorption opacity $\sigma_{\Ab}$, employing various IMEX time stepping schemes, are compared.  Errors for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are: PA2 (triangles pointing left), PA2+ (triangles pointing right), PC2 (asterisk), RKCB2 ($\times$), SSP2332 ($+$), and PARSD (circles).  Black dash-dot reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively.}
  \label{fig:SineWaveDamping}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{figures/SineWaveDiffusionN}
   \caption{Absolute error (cf. Eq.~\eqref{eq:errorNormAbsolute}) for the number density $\cJ$ versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are: PA2 (triangle pointing left), PA2+ (triangle pointing right), PC2 (asterisk), RKCB2 (cross), SSP2332 (plus), and PARSD (circle).  Black dash-dot reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively.}
  \label{fig:SineWaveDiffusionJ}
\end{figure}

@@

In Figures~\ref{fig:SineWaveDiffusionJ} and \ref{fig:SineWaveDiffusionJ} we plot the absolute error, obtained using different values of $\sigma_{\Scatt}$, for various IMEX schemes at $t=t_{\mbox{\tiny end}}$.  
Results for $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively (see figure caption for further details).  
(Scheme PC2 has been shown to work well for this test \cite{radice_etal_2013}, but is included here for comparison with the other IMEX schemes.)
Schemes PARSD, RKCB2, and SSP2332 are accurate for this test, and display third-order accuracy for the number density $\cJ$ and second-oder accuracy for $\cH_{x}$.  
For $\sigma=10^{2}$, the errors do not drop below $10^{-6}$ because of differences between the two-moment model and the diffusion equation used to obtain the analytic solution.  
For larger values of the scattering opacity, the two-moment model agrees better with the diffusion model, and we observe convergence over the entire range of $N$.  
Schemes PA2 and PA2+ do not perform well on this test (for reasons discussed in Section~\ref{sec:imex}).  
For $\sigma_{\Scatt}=10^{2}$, errors in $\cJ$ and $\cH_{x}$ decrease with increasing $N$, but for $\sigma_{\Scatt}=10^{4}$, errors remain constant with increasing $N$ over the entire range.

@@

Here, this test is computed on a three-dimensional Cartesian domain $D=\{\vect{x}\in\bbR^{3}:x^{1}\in[0,2], x^{2}\in[0,2], x^{3}\in[0,2]\}$.  
Because of the symmetry of the problem, and to save computational resources, we only compute the solution in one octant.  
On the inner boundaries, we impose reflecting boundary conditions, while we impose 'homogeneous' boundary conditions on the outer boundary in all three coordinate dimensions; i.e., values for all moments in a boundary element are set equal to the corresponding values in the nearest element just inside $D$.  
Since this test is computed with Cartesian coordinates using a relatively low spatial resolution ($64^{3}$), we have found it necessary to smooth out the opacity over a finite radial extent to avoid numerical artifacts due to a discontinuous absorption opacity.  
Specifically, we use an absorption opacity of the following form
\begin{equation}
  \sigma_{\Ab}(r)=\f{\sigma_{\Ab,0}}{(r/R_{0})^{p}+1}.  
\end{equation}
We set $f_{0}=1$, and compute three versions of this test: one with $\sigma_{\Ab,0}=1$, $R_{0}=1$, and $p=80$ (Test~A), one with $\sigma_{\Ab,0}=10$, $R_{0}=1$, and $p=80$ (Test~B), and one with $\sigma_{\Ab}=10^{3}$, $R_{0}=0.85$, and $p=40$ (Test~C).  
(These values for $R_{0}$ and $p$ result in similar radius for where the optical depth equals $2/3$ in Test~B and Test~C.)
We compute until $t=5$, when the system has reached an approximate steady state.  
We use the IMEX scheme PARSD --- the least compute-intensive of the realizability-preserving IMEX schemes presented here --- in all the tests.  
The main purpose of this test is to compare the results obtained using the different algebraic closures discussed in Section~\ref{sec:algebraicClosure}.

>>>>

To compare the accuracy of the IMEX schemes, we present results from smooth problems in streaming, absorption, and scattering dominated regimes in one spatial dimension.  
For all tests in this subsection, we use third order accurate spatial discretization (polynomials of degree $k=2$) and we employ the maximum entropy closure in the low occupancy limit (i.e., the Minerbo closure).  
We compare results obtained using IMEX schemes proposed here (PA2+ and PD-ARS) with IMEX schemes from Hu et al. \cite{hu_etal_2018} (PA2), McClarren et al. \cite{mcclarren_etal_2008} (PC2), Pareschi \& Russo \cite{pareschiRusso_2005} (SSP2332), and Cavaglieri \& Bewley \cite{cavaglieriBewley2015} (RKCB2).  
In the streaming test, we also include results obtained with second-order and third-order accurate explicit strong stability-preserving Runge-Kutta methods \cite{gottlieb_etal_2001} in the comparison (SSPRK2 and SSPRK3, respectively).  
See \ref{app:butcherTables} for further details.  
The time step is set to $\dt=0.1\times\dx$.

@@

In Figure~\ref{fig:SineWaveStreaming}, the absolute error for the number density $E_{\mbox{\tiny Abs}}^{1}[\cJ_{h}](t=10)$ is plotted versus $N$ (see figure caption for details).  
Errors obtained with SSPRK3 are smallest and decrease as $N^{-3}$ (cf. bottom black dash-dot reference line), as expected for a scheme combining third-order accurate time stepping with third-order accurate spatial discretization.  
For all the other schemes, using second-order accurate explicit time stepping, the error decreases as $N^{-2}$.  
Among the second-order accurate methods, SSP2332 has the smallest error, followed by RKCB2.  
Errors for the remaining schemes (including SSPRK2) are indistinguishable on the plot.  
\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/SineWaveStreaming}
   \caption{Absolute error (cf. Eq.~\eqref{eq:errorNormAbsolute}) versus number of elements $N$ for the streaming sine wave test.  Results employing various time stepping schemes are compared: SSPRK2 (cyan triangles pointing up), SSPRK3 (cyan triangles pointing down), PA2 (red), PA2+ (purple), PC2 (blue), RKCB2 (dark green), SSP2332 (green), and PD-ARS (light red circles).  Black dash-dot reference lines are proportional to $N^{-1}$ (top), $N^{-2}$ (middle), and $N^{-3}$ (bottom), respectively.}
  \label{fig:SineWaveStreaming}
\end{figure}

@@

Figure~\ref{fig:SineWaveDamping} shows convergence results, obtained using different values of $\sigma_{\Ab}$, for various IMEX schemes at $t=t_{\mbox{\tiny end}}$.  
Results for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively (see figure caption for further details).  
All the second-order accurate schemes (PA2, PA2+, RKCB2, and SSP2332) display second-order accurate convergence rates (cf. bottom, black dash-dot reference line).  
For $\sigma_{\Ab}=0.1$, SSP2332 is the most accurate among these schemes, while PA2+ is the most accurate for $\sigma_{\Ab}=10$.  
On the other hand, PC2 and PD-ARS are indistinguishable and display at most first-order accurate convergence, as expected.  
(For $\sigma_{\Ab}=0.1$, PC2 and PD-ARS are the most accurate schemes for $N=8$ and $N=16$.)

\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/SineWaveDamping}
   \caption{Relative error (cf. Eq.~\eqref{eq:errorNormRelative}) versus number of elements for the damping sine wave test.  Results for different values of the absorption opacity $\sigma_{\Ab}$, employing various IMEX time stepping schemes, are compared.  Errors for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are: PA2 (triangles pointing left), PA2+ (triangles pointing right), PC2 (asterisk), RKCB2 ($\times$), SSP2332 ($+$), and PD-ARS (circles).  Black dash-dot reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively.}
  \label{fig:SineWaveDamping}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{figures/SineWaveDiffusionN}
   \caption{Absolute error (cf. Eq.~\eqref{eq:errorNormAbsolute}) for the number density $\cJ$ versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are: PA2 (triangle pointing left), PA2+ (triangle pointing right), PC2 (asterisk), RKCB2 (cross), SSP2332 (plus), and PD-ARS (circle).  Black dash-dot reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively.}
  \label{fig:SineWaveDiffusionJ}
\end{figure}

@@

In Figures~\ref{fig:SineWaveDiffusionJ} and \ref{fig:SineWaveDiffusionJ} we plot the absolute error, obtained using different values of $\sigma_{\Scatt}$, for various IMEX schemes at $t=t_{\mbox{\tiny end}}$.  
Results for $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively (see figure caption for further details).  
(Scheme PC2 has been shown to work well for this test \cite{radice_etal_2013}, but is included here for comparison with the other IMEX schemes.)
Schemes PD-ARS, RKCB2, and SSP2332 are accurate for this test, and display third-order accuracy for the number density $\cJ$ and second-oder accuracy for $\cH_{x}$.  
For $\sigma=10^{2}$, the errors do not drop below $10^{-6}$ because of differences between the two-moment model and the diffusion equation used to obtain the analytic solution.  
For larger values of the scattering opacity, the two-moment model agrees better with the diffusion model, and we observe convergence over the entire range of $N$.  
Schemes PA2 and PA2+ do not perform well on this test (for reasons discussed in Section~\ref{sec:imex}).  
For $\sigma_{\Scatt}=10^{2}$, errors in $\cJ$ and $\cH_{x}$ decrease with increasing $N$, but for $\sigma_{\Scatt}=10^{4}$, errors remain constant with increasing $N$ over the entire range.

@@

Here, this test is computed on a three-dimensional Cartesian domain $D=\{\vect{x}\in\bbR^{3}:x^{1}\in[0,2], x^{2}\in[0,2], x^{3}\in[0,2]\}$.  
Because of the symmetry of the problem, and to save computational resources, we only compute the solution in one octant.  
On the inner boundaries, we impose reflecting boundary conditions, while we impose 'homogeneous' boundary conditions on the outer boundary in all three coordinate dimensions; i.e., values for all moments in a boundary element are set equal to the corresponding values in the nearest element just inside $D$.  
Since this test is computed with Cartesian coordinates using a relatively low spatial resolution ($64^{3}$), we have found it necessary to smooth out the opacity over a finite radial extent to avoid numerical artifacts due to a discontinuous absorption opacity.  
Specifically, we use an absorption opacity of the following form
\begin{equation}
  \sigma_{\Ab}(r)=\f{\sigma_{\Ab,0}}{(r/R_{0})^{p}+1}.  
\end{equation}
We set $f_{0}=1$, and compute three versions of this test: one with $\sigma_{\Ab,0}=1$, $R_{0}=1$, and $p=80$ (Test~A), one with $\sigma_{\Ab,0}=10$, $R_{0}=1$, and $p=80$ (Test~B), and one with $\sigma_{\Ab}=10^{3}$, $R_{0}=0.85$, and $p=40$ (Test~C).  
(These values for $R_{0}$ and $p$ result in similar radius for where the optical depth equals $2/3$ in Test~B and Test~C.)
We compute until $t=5$, when the system has reached an approximate steady state.  
We use the IMEX scheme PD-ARS --- the least compute-intensive of the realizability-preserving IMEX schemes presented here --- in all the tests.  
The main purpose of this test is to compare the results obtained using the different algebraic closures discussed in Section~\ref{sec:algebraicClosure}.
